NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE, PG and PS lipids

### 2017-08-15 Manuscript/manuscript.tex
Add results delivered by Buslaev and some updates in the manuscript

Absolute values of order parameters between Slipids simulations \cite{??}
and experiments for DPPE are shown in Fig. \ref{HGorderParametersPE}.
Glycerol bakcbone order parameters in Slipids are off from experiments,
as systematically observed in this and previous studies \cite{??}.
Alpha order parameter is too low, but beta shows apparent agreement
with experiments. However, the sign of beta order parameter is positive,
in contrast to PC where negative sign was measured. Thus, the the beta order
parameter agrees with experiment with the assumption that its sign is opposite
than for PC.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPE.eps}
  \caption{\label{HGorderParametersPE}
    Order parameters for DPPE headgroup and glycerol
    backbone from simulations with Slipids \cite{??} and experiments
    (DPPE from \cite{seelig76} and E.coliPE from \cite{gally81}).
    Absolute values are shown, because signs are not known experimentally.
  }
  \todo{Experimental signs of the order parameters would be highly useful.} \\
\end{figure}

Comparison between experiments and simulations with CHARMM GUI
parameters for PS lipids in Fig. \ref{HGorderParametersPOPS} suggest that
the model overestimates magnitudes in $\beta$ and g$_3$ order parameters
and forking in $\alpha$.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPOPS.eps}
  \caption{\label{HGorderParametersPOPS}
    Order parameters for POPS (top) and DOPS (bottom) headgroup and glycerol
    backbone from simulations with different models and experiments without CaCl$_2$ 
    (DOPS from \cite{browning80} contains 0.1M of NaCl, POPC:POPS mixture from \cite{roux90}).
    Absolute values are shown for experimental data, because signs are not known.
    Simulations values are -S$_{CH}$
  }
  \todo{More simulation data for lipids with different headgroups to be collected} \\
  \todo{Some simulations contain potassium as counterions, while some sodium.
    All experiments here contain some amount of sodium salt. The best
    ion concentrations for comparison should be figured out.} \\
  \todo{Experimental signs of the order parameters would highly useful.} \\
  \todo{Why there is difference between CHARMM36 simulation results from POPS:POPC mixture and
  pure POPS? Discussion in https://github.com/NMRLipids/NMRlipidsIVotherHGs/issues/1}
\end{figure}

Comparison between experiments and simulations with CHARMM GUI
parameters for PG lipids in Fig. \ref{HGorderParametersPOPG} suggest that
the model overestimates forking in $\gamma$ and $\alpha$ order parameters,
and magnitudes of $\beta$ and $g_1$ order parameters.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPOPG.eps}
  \caption{\label{HGorderParametersPOPG}
    Order parameters for PG headgroup and glycerol backbone from simulations with CHARMM GUI parameters and experiments without CaCl$_2$ 
    (POPG from \cite{borle85} contains 10mM of PIPES, DPPG from \cite{wohlgemuth80} contains 10mM PIPES and 100mM CaCl,
    E.Coli PG results from \cite{gally81}).
    Absolute values are shown, because signs are not known for experimental data.
    Glycerol backbone order parameters are negative in simulations, while $alpha$, $\beta$ and
    smaller $\gamma$ are positive.
  }
  \todo{More simulation data for lipids with different headgroups to be collected}
  \todo{CHARMM GUI simulation contains only counter ions as potassium.
    All experiments here contain some amount of sodium salt. The best
    ion concentrations for comparison should be figured out.}
  \todo{Experimental signs of the order parameters would highly useful.}
\end{figure}

>>>>

\subsection{PE headgroup}
Absolute values of order parameters between Slipids simulations \cite{??}
and experiments for DPPE are shown in Fig. \ref{HGorderParametersPE}.
Glycerol bakcbone order parameters in Slipids are off from experiments,
as systematically observed in this and previous studies \cite{??}.
Alpha order parameter is too low, but beta shows apparent agreement
with experiments. However, the sign of beta order parameter is positive,
in contrast to PC where negative sign was measured. Thus, the the beta order
parameter agrees with experiment with the assumption that its sign is opposite
than for PC.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPE.eps}
  \caption{\label{HGorderParametersPE}
    Order parameters for DPPE headgroup and glycerol
    backbone from simulations with Slipids \cite{??} and experiments
    (DPPE from \cite{seelig76} and E.coliPE from \cite{gally81}).
    Absolute values are shown, because signs are not known experimentally.
  }
  \todo{Experimental signs of the order parameters would be highly useful.} \\
\end{figure}

\subsection{PS headgroup}
Experimental order parameters for bilayers with PS headgroups are 
available in literature only for DOPS  \cite{browning80}. These are shown together
with the results from different simulation models in Fig. \ref{HGorderParametersDOPS}.
The signs are not yet know experimentally.
Before the experimental data becomes available we have denoted 
negative signs for $\beta$ and $\alpha$ segment order parameters,
because this selection of signs is in best agreement with the most simulation
results. With this interpretation the CHARMM36, CHARMM36ua and Slipids
models would be in relatively good agreement with experimental data in
the headgroup region. However, if the experimental signs turn out to
be different than expected, this conclusion will change.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersDOPS.eps}
  \caption{\label{HGorderParametersDOPS}
    Order parameters for DOPS headgroup and glycerol
    backbone from simulations with different models and experiments without CaCl$_2$ 
    Experimental data from \cite{browning80} contains 0.1M of NaCl.
    Signs are not known experimentally. They chosen to give best agreement simulations.
  }
  \todo{Experimental signs of the order parameters would highly useful.} \\
\end{figure}

The glycerol backbone order parameters for Slipids model are, however,
far from experimental values, as previously observed also for PC 
lipids \cite{botan15}. The glycerol backbone structures between PC and
PS lipids simulated with CHARMM36 are compared with the structures
simulated with CHARMM36 in Fig. \ref{glycerol_buslaev}.
The differences in sampled conformation leading to the order parameter
differences are clearly visible in the figure.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/glycerol_buslaev.png}
  \caption{\label{glycerol_buslaev}
    Snapshots overlayed from different simulations for glycerol backbone region
    by Pavel Buslaev.
  }
\end{figure}

%Comparison between experiments and simulations with CHARMM GUI
%parameters for PS lipids in Fig. \ref{HGorderParametersPOPS} suggest that
%the model overestimates magnitudes in $\beta$ and g$_3$ order parameters
%and forking in $\alpha$.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPOPS.eps}
  \caption{\label{HGorderParametersPOPS}
    Order parameters for POPS (top) and DOPS (bottom) headgroup and glycerol
    backbone from simulations with different models and experiments without CaCl$_2$ 
    (DOPS from \cite{browning80} contains 0.1M of NaCl, POPC:POPS mixture from \cite{roux90}).
    Absolute values are shown for experimental data, because signs are not known.
    Simulations values are -S$_{CH}$
  }
  \todo{More simulation data for lipids with different headgroups to be collected} \\
  \todo{Some simulations contain potassium as counterions, while some sodium.
    All experiments here contain some amount of sodium salt. The best
    ion concentrations for comparison should be figured out.} \\
  \todo{Experimental signs of the order parameters would highly useful.} \\
  \todo{Why there is difference between CHARMM36 simulation results from POPS:POPC mixture and
  pure POPS? Discussion in https://github.com/NMRLipids/NMRlipidsIVotherHGs/issues/1}
\end{figure}

\subsection{PG headgroup}
Comparison between experiments and simulations with CHARMM GUI
parameters for PG lipids in Fig. \ref{HGorderParametersPOPG} suggest that
the model overestimates forking in $\gamma$ and $\alpha$ order parameters,
and magnitudes of $\beta$ and $g_1$ order parameters.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPOPG.eps}
  \caption{\label{HGorderParametersPOPG}
    Order parameters for PG headgroup and glycerol backbone from simulations with CHARMM GUI parameters and experiments without CaCl$_2$ 
    (POPG from \cite{borle85} contains 10mM of PIPES, DPPG from \cite{wohlgemuth80} contains 10mM PIPES and 100mM CaCl,
    E.Coli PG results from \cite{gally81}).
    Absolute values are shown, because signs are not known for experimental data.
    Glycerol backbone order parameters are negative in simulations, while $alpha$, $\beta$ and
    smaller $\gamma$ are positive.
  }
  \todo{More simulation data for lipids with different headgroups to be collected}
  \todo{CHARMM GUI simulation contains only counter ions as potassium.
    All experiments here contain some amount of sodium salt. The best
    ion concentrations for comparison should be figured out.}
  \todo{Experimental signs of the order parameters would highly useful.}
\end{figure}
